---
title: "Mapping of URPs onto an ensemble of PDB structures"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(curl)
library(progress)
```


Generated from main analysis script
```{r}
CombinedURPs <- read.table("../MainAnalysis/URPsForMapping.txt", stringsAsFactors=FALSE, sep = "\t", header = TRUE)
```

Annotating information for individual crosslinked residues using the EBI resource SIFTS
```{r}
#downloaded from https://www.ebi.ac.uk/pdbe/docs/sifts/quick.html on 17/3/21
UniprotSegments <- read.delim(file = "Data/SIFTS/uniprot_segments_observed.tsv", comment.char="#", stringsAsFactors = FALSE)

#separating A/B residues (are different in interlinks)
CombinedURPs <- CombinedURPs %>% 
  separate(URP_Key, sep = "_", into = c("Res1", "Res2"), remove = FALSE) %>% 
  separate(Res1, sep = "-", into = c("accessionA", "modposA"), remove = FALSE) %>% 
  separate(Res2, sep = "-", into = c("accessionB", "modposB"), remove = FALSE)


AllA <- CombinedURPs %>% 
  select(Res1, accessionA, modposA) %>% 
  rename(Accession = accessionA, modPos = modposA, Residue = Res1)%>% 
  mutate(modPos = as.numeric(modPos))

AllB <- CombinedURPs %>% 
  select(Res2, accessionB, modposB) %>% 
  rename(Accession = accessionB, modPos = modposB, Residue = Res2) %>% 
  mutate(modPos = as.numeric(modPos))

#merging A/B residues into single list
All <- bind_rows(AllA, AllB) %>% 
  mutate(Position = modPos + 1) %>% 
  # adding +1 to position due to XLDB numbering being -1 compared to PDB - consistency to allow matching
  select(Accession, Position, Residue) %>% 
  distinct()
remove(AllA, AllB)

#merge residues with PDB info
AllPDB <- left_join(All, UniprotSegments, by = c("Accession" = "SP_PRIMARY"))

remove(All)
ResolvedPDB <- AllPDB %>% 
  filter((Position >= SP_BEG) &(Position <= SP_END)) %>% 
  mutate(shift = as.numeric(PDB_BEG) - as.numeric(SP_BEG)) %>% 
  mutate(ResInStructure = Position + shift)

ResolvedPDBSuucinct <- ResolvedPDB %>% 
  select(Residue, PDB, CHAIN, ResInStructure) %>% 
  rename(PDB1 = PDB, CHAIN1 = CHAIN, ResInStructure1 = ResInStructure)

CombinedURPs <- left_join(CombinedURPs, ResolvedPDBSuucinct, by = c("Res1" = "Residue"))

ResolvedPDBSuucinct <- ResolvedPDBSuucinct %>% 
   rename(PDB2 = PDB1, CHAIN2 = CHAIN1, ResInStructure2 = ResInStructure1)

CombinedURPs <- left_join(CombinedURPs, ResolvedPDBSuucinct, by = c("Res2" = "Residue"))

MappedURPs <- CombinedURPs %>% 
  filter(PDB1 == PDB2) %>% 
  distinct() %>% 
  mutate(ID = row_number())

PDBStructures <- ResolvedPDB %>% 
  select(PDB) %>% 
  distinct()

```

```{r}
StructuresWithMappedXLs <- MappedURPs %>% 
  select(PDB1) %>% 
  distinct()
```


```{r}
PDBStructureNumberMappings <- MappedURPs %>% 
  group_by(PDB1) %>% 
  summarise(NumMappings  = n())
```

```{r}
UniprotSeqs <- read.delim(file = "../../Crosslinking/MainAnalysis/Data/UniProt/uniprot-filtered-reviewed_yes+AND+organism__Saccharomyces+cerevisiae+(st--.tab")
```


Generating a table of URPs for each PDB structure with crosslinks
```{r}
dir.create("Xwalk_Inputs")

for(i in 1:NROW(StructuresWithMappedXLs)) {


Structure <- StructuresWithMappedXLs$PDB[i]

CrosslinksToMap <- MappedURPs %>% 
  mutate(AminoAcid1 = "LYS") %>% 
  mutate(AminoAcid2 = "LYS") %>% 
  filter(PDB1 == Structure ) 

CrosslinksToMapWithURP <- CrosslinksToMap %>% 
  mutate(filename = paste0(Structure, ".pdb")) %>% 
  mutate(LinkSite1 = paste0(AminoAcid1,"-", ResInStructure1,"-", CHAIN1, "-CA")) %>% 
  mutate(LinkSite2 = paste0(AminoAcid2,"-", ResInStructure2,"-", CHAIN2, "-CA")) %>% 
  select(URP_Key, filename,LinkSite1, LinkSite2) %>% 
  distinct()

RandomCrosslinksWithURP <- data.frame(matrix(NA, ncol=4, nrow=NROW(CrosslinksToMap)))
colnames(RandomCrosslinksWithURP) <- c("URP_Key", "filename", "LinkSite1", "LinkSite2" )

for(j in 1:NROW(CrosslinksToMap)) {

    ResiduesK1 <- str_locate_all(UniprotSeqs$Sequence[UniprotSeqs$Entry ==  CrosslinksToMap$Accession1[j]], "K")[[1]][,1]

  Random_K1 <-   sample(ResiduesK1)[1]
  
  ResiduesK2 <- str_locate_all(UniprotSeqs$Sequence[UniprotSeqs$Entry ==  CrosslinksToMap$Accession2[j]], "K")[[1]][,1]
  Random_K2 <- sample(ResiduesK2)[1]
  
  if(Random_K1 == Random_K2) {
      Random_K2 <- sample(ResiduesK2)[1]
  }

RandomCrosslinksWithURP$filename[j] <- CrosslinksToMapWithURP$filename[j]
RandomCrosslinksWithURP$LinkSite1[j] <- paste0(CrosslinksToMap$AminoAcid1[j],"-",Random_K1 ,"-", CrosslinksToMap$CHAIN1[j], "-CA")
RandomCrosslinksWithURP$LinkSite2[j] <- paste0(CrosslinksToMap$AminoAcid2[j],"-",Random_K2 ,"-", CrosslinksToMap$CHAIN2[j], "-CA")
RandomCrosslinksWithURP$URP_Key[j] <- paste0(CrosslinksToMap$Accession1[j],"-",Random_K1 ,"_", CrosslinksToMap$Accession2[j], "-", Random_K2)

}



RandomCrosslinksToMap <- RandomCrosslinksWithURP %>% 
    select(filename,LinkSite1, LinkSite2) 

CrosslinksToMap <- CrosslinksToMapWithURP %>% 
    select(filename,LinkSite1, LinkSite2) 

write.table(RandomCrosslinksWithURP, file = paste0("Xwalk_Inputs/", Structure, "_Random_XLs_Xwalk-Input_withURPCode.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)



write.table(RandomCrosslinksToMap, file = paste0("Xwalk_Inputs/", Structure, "_Random_XLs_Xwalk-Input.txt"), sep = "\t", quote = FALSE, col.names = FALSE, row.names = TRUE)
write.table(RandomCrosslinksWithURP, file = paste0("Xwalk_Inputs/", Structure, "_Random_XLs_Xwalk-Input_withURPCode.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)

write.table(CrosslinksToMap, file = paste0("Xwalk_Inputs/", Structure, "_XLs_Xwalk-Input.txt"), sep = "\t", quote = FALSE, col.names = FALSE, row.names = TRUE)
write.table(CrosslinksToMapWithURP, file = paste0("Xwalk_Inputs/", Structure, "_XLs_Xwalk-Input_withURPCode.txt"), sep = "\t", quote = FALSE, col.names = TRUE, row.names = FALSE)

print(i)
}
```


```{r}

PDBs <- unique(ResolvedPDB$PDB)

for(i in 1:length(PDBs)) {
  Structure =PDBs[i]

  
  PDB_Proteins_Chains <- ResolvedPDB %>% 
  filter(PDB == Structure) %>% 
  select(PDB, Accession, CHAIN) %>% 
  distinct()

AllPairTable <- data.frame(matrix(NA, ncol=0, nrow=0))

for(j in 1:NROW(PDB_Proteins_Chains)) {

  ResiduesK <- str_locate_all(UniprotSeqs$Sequence[UniprotSeqs$Entry ==  PDB_Proteins_Chains$Accession[j]], "K")[[1]][,1]
  
  AllLysines <- as.data.frame(ResiduesK)
  AllLysines$Accession = PDB_Proteins_Chains$Accession[j]
  AllLysines$Chain = PDB_Proteins_Chains$CHAIN[j]
  AllLysines$Type = "AllPairs"
  
    Check <- AllPDB %>% 
    filter(Accession == PDB_Proteins_Chains$Accession[j]) %>% 
    filter(PDB == Structure) %>% 
    select(Accession, CHAIN, PDB, RES_BEG:SP_END )
  
  AllLysines <- left_join(AllLysines, Check, by = c("Accession", "Chain" = "CHAIN"))
  
  
  AllLysines <- AllLysines %>% 
  mutate(shift = as.numeric(PDB_BEG) - as.numeric(SP_BEG)) %>% 
  mutate(ResKInStructure = ResiduesK + shift)

  FilteredLysines <- AllLysines %>%   
    mutate(Resolved = if_else(((ResiduesK >= SP_BEG) &(ResiduesK <= SP_END)), TRUE, FALSE)) %>% 
    group_by(PDB, Accession, Chain, ResiduesK, Type, ResKInStructure) %>% 
    summarise(ResolvedAny = any(Resolved)) %>% 
    filter(ResolvedAny == TRUE)
    
  AllPairTable <- bind_rows(AllPairTable, FilteredLysines)
  }

AllPairTable <- AllPairTable %>%
  mutate(ResidueKey = paste0("LYS-", ResKInStructure, "-", Chain, "-CA")) %>% 
  select(ResidueKey, Type)

AllCombinations = full_join(AllPairTable, AllPairTable, by = "Type")

AllCombinations <- AllCombinations %>%
  mutate(filename = Structure) %>% 
  mutate(URPKey = if_else(ResidueKey.x < ResidueKey.y, paste0(ResidueKey.x, "_", ResidueKey.y), paste0(ResidueKey.y, "_", ResidueKey.x))) %>% 
  filter(ResidueKey.x != ResidueKey.y) %>% 
  select(filename, URPKey) %>% 
  distinct() %>% 
  separate(URPKey, into = c("LinkSite1", "LinkSite2"), sep = "_")  
  

write.table(AllCombinations, file = paste0("Xwalk_Inputs/", Structure, "_AllPairs_Xwalk-Input.txt"), sep = "\t", quote = FALSE, col.names = FALSE, row.names = TRUE)

command <- paste0("java -Xmx1024m -cp /Users/tarabartolec/Downloads/xwalk_v0.6/bin Xwalk -infile PDBs/", Structure, ".pdb -out Xwalk_Outputs/", Structure, "_AllPairs.xwalkoutput.tsv -dist Xwalk_Inputs/", Structure, "_AllPairs_Xwalk-Input.txt -max 50 -bb -f -euc" )

system(command)

print(i)
}


```


Downloading the pdb files from the RCSB PDB website
```{r eval=FALSE}

dir.create("PDBs")

NROW(StructuresWithMappedXLs)
for (i in 1:NROW(StructuresWithMappedXLs)) {

pdb = StructuresWithMappedXLs$PDB[i]
url = paste0("https://files.rcsb.org/download/", pdb, ".pdb")
name = paste0("PDBs/", pdb, ".pdb")

req <- curl_fetch_disk(url, name)
req$status_code

if(!(req$status_code %in% c(200, 202))){
    unlink(name)
  print(paste("Number: " , i, " ", name, "removed"))
} 
}
```

Using xwalk to map URPs on and calculate euclidean distance
```{r, eval=FALSE}
dir.create("Xwalk_Outputs")


for(i in 1:NROW(StructuresWithMappedXLs)) {

pdb = StructuresWithMappedXLs$PDB[i]

NumForPDB <- PDBStructureNumberMappings %>% 
  filter(PDB1 == pdb)

NumOfCrosslinkMappings = NumForPDB$NumMappings[1]

if(NumOfCrosslinkMappings < 10000) {
command <- paste0("java -Xmx1024m -cp /Users/tarabartolec/Downloads/xwalk_v0.6/bin Xwalk -infile PDBs/", pdb, ".pdb -out Xwalk_Outputs/", pdb, ".xwalkoutput.tsv -dist Xwalk_Inputs/", pdb, "_XLs_Xwalk-Input.txt -max 50 -bb -f -euc" )

command2 <- paste0("java -Xmx1024m -cp /Users/tarabartolec/Downloads/xwalk_v0.6/bin Xwalk -infile PDBs/", pdb, ".pdb -out Xwalk_Outputs/", pdb, "_Random.xwalkoutput.tsv -dist Xwalk_Inputs/", pdb, "_Random_XLs_Xwalk-Input.txt -max 50 -bb -f -euc" )

system(command)
system(command2)

print(i)
}
}
```

```{r, warning=FALSE, eval=FALSE}

URPsMappedToStructure <- data.frame("", "", "", "", "", "", "")
colnames(URPsMappedToStructure) <- c("Index", "filename", "LinkSite1", "LinkSite2", "DistAA", "CAEuclideanDistance", "URP_Key")

for (i in 1:NROW(StructuresWithMappedXLs)) {

pdb = StructuresWithMappedXLs$PDB[i]
xwalkoutputname =  paste0("Xwalk_Outputs/", pdb, ".xwalkoutput.tsv") 
xwalkinputURPname = paste0("Xwalk_Inputs/", pdb, "_XLs_Xwalk-Input_withURPCode.txt")

xwalkoutput <- NA
xwalkoutput <- try(read.delim(file = xwalkoutputname, stringsAsFactors = FALSE, header = FALSE))
try({colnames(xwalkoutput) <- c("Index", "filename", "LinkSite1", "LinkSite2", "DistAA", "CAEuclideanDistance")})
try({xwalkoutput <- xwalkoutput %>% 
  select(1:6)})

xwalkinputURP <- read.delim(file = xwalkinputURPname, stringsAsFactors = FALSE, header = TRUE)

try(xwalkoutput <- left_join(xwalkoutput, xwalkinputURP))

URPsMappedToStructure <- rbind(URPsMappedToStructure, xwalkoutput)

print(i)
}

write.table(URPsMappedToStructure, file = "URPsMappedToStructure.txt", row.names = FALSE, quote = FALSE, sep = "\t")

```


#doing the same for the random subset
```{r warning=FALSE, eval=FALSE}
RandomURPsMappedToStructure <- data.frame("", "", "", "", "", "", "")
colnames(RandomURPsMappedToStructure) <- c("Index", "filename", "LinkSite1", "LinkSite2", "DistAA", "CAEuclideanDistance", "URP_Key")

for (i in 1:NROW(StructuresWithMappedXLs)) {

pdb = StructuresWithMappedXLs$PDB[i]
xwalkoutputname =  paste0("Xwalk_Outputs/", pdb, "_Random.xwalkoutput.tsv") 
xwalkinputURPname = paste0("Xwalk_Inputs/", pdb, "_Random_XLs_Xwalk-Input_withURPCode.txt")

xwalkoutput <- NA
xwalkoutput <- try(read.delim(file = xwalkoutputname, stringsAsFactors = FALSE, header = FALSE))
try({colnames(xwalkoutput) <- c("Index", "filename", "LinkSite1", "LinkSite2", "DistAA", "CAEuclideanDistance")})
try({xwalkoutput <- xwalkoutput %>% 
  select(1:6)})

xwalkinputURP <- read.delim(file = xwalkinputURPname, stringsAsFactors = FALSE, header = TRUE)

try(xwalkoutput <- left_join(xwalkoutput, xwalkinputURP))

RandomURPsMappedToStructure <- rbind(RandomURPsMappedToStructure, xwalkoutput)
}

write.table(RandomURPsMappedToStructure, file = "RandomURPsMappedToStructure.txt", row.names = FALSE, quote = FALSE, sep = "\t")
```


Summarising distances across structures by taking the minimum distance imposed for a URP on any structure
```{r}
#A pre-compiled file to save time on reanalysis
URPsMappedToStructure <-  read.delim("URPsMappedToStructure.txt")
RandomURPsMappedToStructure <-  read.delim("RandomURPsMappedToStructure.txt")

URPsMappedToStructure$Category = "Crosslinks"

RandomURPsMappedToStructure$Category = "Random"

URPsMappedToStructure <- bind_rows(URPsMappedToStructure, RandomURPsMappedToStructure)


URPsMappedToStructureSummary <- URPsMappedToStructure %>% 
  filter(!(str_detect(Index, "Error")))  %>% 
  mutate(CAEuclideanDistance = as.numeric(CAEuclideanDistance)) %>% 
  group_by(URP_Key, Category) %>% 
  summarise(minDist = min(CAEuclideanDistance), maxDist = max(CAEuclideanDistance), numMappings = n(), numStructures = n_distinct(filename))  %>% 
  mutate(RangeOfDistance = maxDist - minDist) %>% 
  filter(!is.na(RangeOfDistance))


StructureSummary <- URPsMappedToStructure %>% 
   filter(!(str_detect(Index, "Error"))) %>% 
  group_by(filename, URP_Key) %>% 
  summarise(MinDist = min(CAEuclideanDistance)) %>% 
  mutate(Valid = MinDist <= 42)

StructureSummary2 <- StructureSummary %>% 
  filter(filename != "") %>% 
  group_by(filename) %>% 
  summarise(NumURPs = n_distinct(URP_Key), NumValid = sum(Valid)) %>% 
  mutate(PercentValid = 100*(NumValid/NumURPs) )


NROW(unique(URPsMappedToStructure$filename))
NROW(unique(URPsMappedToStructureSummary$URP_Key))

write_tsv(URPsMappedToStructure, file = "URPsMappedToStructure.txt" )


ggplot(URPsMappedToStructureSummary, aes(x = minDist, fill = Category)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 42) +
   scale_fill_manual(values = c("blue", "#bababa") )+
  #ggtitle(paste0( NROW(URPsMappedToStructureSummary), " URPs mapped across ",  NROW(unique(URPsMappedToStructure$filename)) ," pdb structures")) +
  xlab("Min. Ca-Ca euclidean distance of URP\nin any PDB file and subunit configuration") +
    theme_bw() +
   theme(legend.position = "none", 
          text=element_text(size=10,  family="Arial"), 
          panel.border = element_rect(fill=NA, colour = "black", size=1),
          panel.grid.minor = element_blank())


ggsave(filename = "URPsOnStructure.pdf", device = cairo_pdf, 
       width = 2.5, height = 1.65, units = "in")
```






